---
title: "Loading ibutton data"
author: "Andrew MacDonald"
date: "`r Sys.Date()`"
vignette: >
  %\VignetteIndexEntry{reading and tidying ibutton output}
  %\VignetteEngine{knitr::knitr}
  %\usepackage[utf8]{inputenc}
---

# Loading and cleaning ibutton data
**Andrew MacDonald, Oct 2014**

[ibutton](http://en.wikipedia.org/wiki/1-Wire) temperature loggers are frequently used to monitor temperature in nature.  They are tough and versatile, and have been used in many applications.

The data that ibuttons produce requires a bit of cleaning before use in data analysis. `ibuttonr` automates the process of reading and cleaning your ibutton data.

## Loading data

`ibuttonr` assumes that you have all your ibutton data in a single folder, stored as .csv files. For advice on getting your data out of an ibutton and into a csv, see the [product website](http://www.maximintegrated.com/en/app-notes/index.mvp/id/3358).  

Once you have all your data in a single folder, simply point the function `read.ibutton.folder` at it.  Here is how you would do so if the working directory is set to your ibutton data folder (`.`):

```{r}
library(ibuttonr)
ibutton.data <- read_ibutton_folder(".")
str(ibutton.data)
```

**NOTE** this function assumes that the directory has *only* ibutton csv files inside. 

if your working directory is set elsewhere, you can use a relative path to the folder, such as `ibuttondata/`.

## Checking the data

### Stopped ibuttons
Ibuttons frequently fail, but which ones? you could just check the number of rows in each, and pick
the ones which are suspiciously short, indicating that the ibutton stopped
recording temperature. This function automates this process to make it a bit
more objective: it points out ones which recorded less than the median number
of datapoints for the experiment. It assumes that all the ibuttons were
supposed to run for equal amounts of time.
```{reval=FALSE}

id_broken(ibutton.data)
## get the lengths for your own use:
sapply(ibutton.data,nrow)
```

### Duplicate registration numbers
sometimes, fieldworkers record the data incorrectly -- for example, one
common mistake is to save data from the same ibutton twice with different
filenames. However, each ibutton has a unique registration number. check
for any number >1 in this table to identify this error. Additionally, if you
recorded the registration numbers (written on the back of the ibuttons) you
could use this to get them from the datafiles themselves
```{r }

table(get_registration_numbers("."))

```

correct the dates and make dataframe
Now that the data is checked, this function takes the list of ibuttons and
combines them together. It also reformats the "Date.Time" variable, into a
format that R recognizes as a date and time.

```{r results='asis'}
ibutton.dataframe <- ibuttons_to_data_frame(ibutton.data)
head(ibutton.dataframe)
knitr::kable(summary(ibutton.dataframe))
```

```{r}
## Here are some ways to graph this data:
with(subset(ibutton.dataframe,ibutton=="10a"),plot(Date.Time,Value,type="n"))
for(i in levels(ibutton.dataframe$ibutton)){
x <- which(ibutton.dataframe$ibutton==i)
lines(ibutton.dataframe$Date.Time[x],ibutton.dataframe$Value[x])
}
###